First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-08-01 ms.tex
bolded change from previous commit
Addressing minor comments from referee

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}.
Furthermore, \citet{Pober:2015} and \citet{Greig:2016} were able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}, 
while others are investigating the spectral structure of point sources themselves \cite[e.g.][]{Offringa:2016}.
The EoR window is still expected to be preserved above the wedge, and 
several analysis pipelines are under active development \citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2015b, Dillon:2013, Trott:2014}.

@@

The data for this work includes two sets. First we use a set of 94 two-minute high band observations of 
the EoR0 field from August 23, 2013, which were taken early in telescope operations and 
found to be particularly well behaved. The MWA EoR collaboration has designated this set 
as a ``golden data set'', which is used to build analysis tools and compare early results. All 
techniques demonstrated here will use this golden data set \cite[e.g.][]{Jacobs:2016, Trott:2016, Pober:2016, Dillon:2015}. In Section~\ref{sec:deep} we 
will use a larger data set to move toward a deep integration. This data set consists of all 
EoR0 high band observations beginning August 23, 2013 (including the golden data set), 
and concluding when the field was no longer accessible for the season: when the sun was 
above the horizon for most or all of the field's transit. The final observation of this set was 
on November 29, 2013. The individual snapshots are 112 seconds of data at 0.5 second, 
40 kHz resolution. In total we analyze 2,780 snapshots, or about 86.5 hours of data.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
80~kHz frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citet{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[Hazelton et al. 2016, in prep.]{Barry:2016}.

@@

Under these assumptions we can express the measured, uncalibrated visibility for an 
antenna pair $(i,j)$ as
\begin{equation}
V'_{ij}(\nu) \approx g_i(\nu)g^*_j(\nu)V_{ij}(\nu),
\end{equation}
where $g_i(\nu)$ and $g_j(\nu)$ are the complex gains for antennas $i$ and $j$ 
respectively. Because we are treating polarizations independently, we allow the antenna 
subscript to run over both East-West (E-W) and North-South (N-S) polarizations.

@@

It has been shown that subtracting a foreground model strictly within the main lobe of the 
primary beam will not be sufficient to suppress the power spectrum wedge and unlock the 
EoR window \citep{Thyagarajan:2015b, Thyagarajan:2015, Pober:2016}. 
We are in the process of repeating the model building described above using additional
MWA observations pointed away from our field of interest.
Until that work is complete we supplement the extent of 
our point source model using additional catalogs. We accomplish this through a hierarchical 
catalog pulling from our early KGS catalog, the MWA Commissioning Survey\footnote{The 
more complete GLEAM Survey \citep[][Hurley-Walker et al., in collaboration review]
{Wayth:2015} was not available at the time of analysis.} \citep[MWACS, ][]
{Hurley-Walker:2014}, the Culgoora catalog \citep{Slee:1977}, and the Molonglo Reference Catalog 
\citep[MRC, ][]{Large:1981}. Source flux densities are prioritized in this order based on our 
confidence in their predicted flux density at 182 MHz. We first cluster the source lists to avoid 
redundant sources, using a 3.5 arcmin neighborhood radius, then select the flux density from the 
highest priority catalog. Spectral indices were obtained from the MWACS and Culgoora 
catalogs when available, otherwise a two-point spectral index was estimated for Culgoora-
MRC matches or MRC-SUMSS matches. All other sources were given a spectral index of 
$-0.8$. The spectral index was used to extrapolate the flux density from the catalog frequency to 
182 MHz, but a uniform spectral index of $-0.8$ is used within our band when forming 
model visibilities for calibration and foreground subtraction.

@@

Future iterations of this model will contain a spectral index 
and multiple polarization components.
Lenc, et al. (in review) demonstrated the presence of strong polarized large-scale structures 
in the EoR0 field at 154 MHz with varying Faraday depths. They observed rotations in the
Stokes Q--U plane due to ionospheric conditions, which will need to be taken into account
in our future diffuse models.

@@

The source catalog used with RTS was a combined cross-match created with the Positional 
Update and Matching Algorithm (PUMA, J.~L.~B.~Line et al. 2016, in review). Full details of PUMA and 
this cross-matched catalog can be found in Line et al. 2016, but we include a brief 
description here. PUMA is designed to cross-match low radio-frequency ($\lesssim 1\,$GHz) 
catalogs. It utilizes a Bayesian probabilistic positional cross-match approach based on~
\citet{Budavari:2008}, combined with criteria based on catalog resolution and fitting the 
spectral energy distribution to a power-law. These criteria allow for surveys with differing 
resolutions, which can introduce confusing sources into the matched process.
Primarily the source catalog was based on the MWA Commissioning Survey
\citep[MWACS, ][]{Hurley-Walker:2014}, which was cross-matched to the $74\,$MHz Very 
Large Array Low Frequency Sky Survey redux~\citep[VLSSr, ][]{Lane:2012}, the Molonglo 
Reference Catalogue~\citep[MRC, ][]{Large:1981}, the $843\,$MHz Sydney University 
Molonglo Sky Survey~\citep[SUMSS,][]{Mauch:2003}; and the $1.4\,$GHz NRAO VLA Sky 
Survey~\citep[NVSS,][]{Condon:1998}. MWACS covers approximately 
$20.5^h < \textrm{RA} < 8.5^h$,  $-ˆ’58^\circ < \delta < -ˆ’14^\circ$, so to complete the sky 
coverage needed, MRC was used as a base catalogue outside of the MWACS coverage and 
was matched to VLSSr, SUMSS and NVSS.

@@

\textbf{Diffuse emission model.} The diffuse model here contained no spectral information
and represented only the Stokes I polarization. Nevertheless subtracting this model removed
90\% of the residual power in the foreground wedge. Adding a spectral index and multiple
polarization components to the model will improve this subtraction even further.
As Lenc, et al. 2016 (in review) demonstrated, substantial polarized diffuse structure
exists in the EoR0 field. The contributions to Stokes Q and U depend on the ionospheric
conditions, which will be accounted for in future analyses to appropriately model this
emission. 
In addition the diffuse model will be extended to an all-sky model. It is evident from the
jackknives shown in Figure~\ref{fig:jackknife} that strong emission near the horizon is
leaking into our power spectra during off-zenith observations. By modeling and subtracting
this emission we may be able to recover these observations in future analyses.

\textbf{Primary beam model.} In order to properly model the instrument response to the sky
we also need an accurate primary beam model. This is especially important for EoR
power spectrum measurements because the (often bright) near-horizon emission resides
at the edge of the foreground wedge 
\citep{Pober:2016, Thyagarajan:2015, Thyagarajan:2015b}, but the low elevation beam
can be difficult to model at the necessary precision \citep{Thyagarajan:2016}. 
\citet{Asad:2016} demonstrated the need for a polarized primary beam model in order
to mitigate the risk of diffuse polarized emission leaking into Stokes I, conflating with a 
potential cosmological measurement. The MWA
primary beam model is continuously improving to incorporate mutual coupling between the 
phased dipoles, and embedded element patterns for each of the 32 dipoles on each antenna 
\citep{Sutinjo:2015}. Other methods are also being pursued to directly measure the full beam 
response in situ using the ORBCOMM satellite constellation \citep{Neben:2015}, and 
drone-flown calibrator sources (Jacobs, et al. in prep). The latter method is similar to
the strategy \citet{Virone:2014} used to verify SKA beam patterns.

>>>>

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}.
\textbf{Furthermore,} \citet{Pober:2015} \textbf{and \citet{Greig:2016} were} able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}, 
while others are investigating the spectral structure of point sources themselves \cite[e.g.][]{Offringa:2016}.
The EoR window is still expected to be preserved above the wedge, and 
several analysis pipelines are under active development 
\textbf{to exploit this foreground isolation}
\citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2015b, Dillon:2013, Trott:2014}.

@@

The data for this work includes two sets. First we use a set of 94 two-minute high band observations of 
the EoR0 field from August 23, 2013, which were taken early in telescope operations and 
found to be particularly well behaved. The MWA EoR collaboration has designated this set 
as a ``golden data set'', which is used to build analysis tools and compare early results 
\textbf{\cite[e.g.][]{Jacobs:2016, Trott:2016, Pober:2016, Dillon:2015}.} All 
techniques demonstrated here will use this golden data set. In Section~\ref{sec:deep} we 
will use a larger data set to move toward a deep integration. This data set consists of all 
EoR0 high band observations beginning August 23, 2013 (including the golden data set), 
and concluding when the field was no longer accessible for the season: when the sun was 
above the horizon for most or all of the field's transit. The final observation of this set was 
on November 29, 2013. The individual snapshots are 112 seconds of data at 0.5 second, 
40 kHz resolution. In total we analyze 2,780 snapshots, or about 86.5 hours of data.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
80~kHz frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citet{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[\textbf{B.~J.}~Hazelton et al. 2016, in prep.]{Barry:2016}.

@@

Under these assumptions we can express the measured, uncalibrated visibility for an 
antenna pair $(i,j)$ as
\begin{equation}
V'_{ij}(\nu) \approx g_i(\nu)g^*_j(\nu)V_{ij}(\nu),
\end{equation}
where $g_i(\nu)$ and $g_j(\nu)$ are the complex gains for antennas $i$ and $j$ 
respectively,
\textbf{and $V_{ij}(\nu)$ is the true visibility which we aim to recover through calibration.}
Because we are treating polarizations independently, we allow the antenna 
subscript to run over both East-West (E-W) and North-South (N-S) polarizations.

@@

It has been shown that subtracting a foreground model strictly within the main lobe of the 
primary beam will not be sufficient to suppress the power spectrum wedge and unlock the 
EoR window \citep{Thyagarajan:2015b, Thyagarajan:2015, Pober:2016}. 
We are in the process of repeating the model building described above using additional
MWA observations pointed away from our field of interest.
Until that work is complete we supplement the extent of 
our point source model using additional catalogs. We accomplish this through a hierarchical 
catalog pulling from our early KGS catalog, the MWA Commissioning Survey\footnote{The 
more complete GLEAM Survey \citep[][\textbf{N.}~Hurley-Walker et al., in review]
{Wayth:2015} was not available at the time of analysis.} \citep[MWACS, ][]
{Hurley-Walker:2014}, the Culgoora catalog \citep{Slee:1977}, and the Molonglo Reference Catalog 
\citep[MRC, ][]{Large:1981}. Source flux densities are prioritized in this order based on our 
confidence in their predicted flux density at 182 MHz. We first cluster the source lists to avoid 
redundant sources, using a 3.5 arcmin neighborhood radius, then select the flux density from the 
highest priority catalog. Spectral indices were obtained from the MWACS and Culgoora 
catalogs when available, otherwise a two-point spectral index was estimated for Culgoora-
MRC matches or MRC-SUMSS matches. All other sources were given a spectral index of 
$-0.8$. The spectral index was used to extrapolate the flux density from the catalog frequency to 
182 MHz, but a uniform spectral index of $-0.8$ is used within our band when forming 
model visibilities for calibration and foreground subtraction.

@@

Future iterations of this model will contain a spectral index 
and multiple polarization components.
\textbf{E.}~Lenc, et al. (in review) demonstrated the presence of strong polarized large-scale structures 
in the EoR0 field at 154 MHz with varying Faraday depths. They observed rotations in the
Stokes Q--U plane due to ionospheric conditions, which will need to be taken into account
in our future diffuse models.

@@

The source catalog used with RTS was a combined cross-match created with the Positional 
Update and Matching Algorithm (PUMA, J.~L.~B.~Line et al. 2016, in review). Full details of PUMA and 
this cross-matched catalog can be found in Line et al. (2016), but we include a brief 
description here. PUMA is designed to cross-match low radio-frequency ($\lesssim 1\,$GHz) 
catalogs. It utilizes a Bayesian probabilistic positional cross-match approach based on~
\citet{Budavari:2008}, combined with criteria based on catalog resolution and fitting the 
spectral energy distribution to a power-law. These criteria allow for surveys with differing 
resolutions, which can introduce confusing sources into the matched process.
Primarily the source catalog was based on the MWA Commissioning Survey
\citep[MWACS, ][]{Hurley-Walker:2014}, which was cross-matched to the $74\,$MHz Very 
Large Array Low Frequency Sky Survey redux~\citep[VLSSr, ][]{Lane:2012}, the Molonglo 
Reference Catalogue~\citep[MRC, ][]{Large:1981}, the $843\,$MHz Sydney University 
Molonglo Sky Survey~\citep[SUMSS,][]{Mauch:2003}; and the $1.4\,$GHz NRAO VLA Sky 
Survey~\citep[NVSS,][]{Condon:1998}. MWACS covers approximately 
$20.5^h < \textrm{RA} < 8.5^h$,  $-ˆ’58^\circ < \delta < -ˆ’14^\circ$, so to complete the sky 
coverage needed, MRC was used as a base catalogue outside of the MWACS coverage and 
was matched to VLSSr, SUMSS and NVSS.

@@

\textbf{Diffuse emission model.} The diffuse model here contained no spectral information
and represented only the Stokes I polarization. Nevertheless subtracting this model removed
90\% of the residual power in the foreground wedge. Adding a spectral index and multiple
polarization components to the model will improve this subtraction even further.
As \textbf{E.}~Lenc, et al. 2016 (in review) demonstrated, substantial polarized diffuse structure
exists in the EoR0 field. The contributions to Stokes Q and U depend on the ionospheric
conditions, which will be accounted for in future analyses to appropriately model this
emission. 
In addition the diffuse model will be extended to an all-sky model. It is evident from the
jackknives shown in Figure~\ref{fig:jackknife} that strong emission near the horizon is
leaking into our power spectra during off-zenith observations. By modeling and subtracting
this emission we may be able to recover these observations in future analyses.

\textbf{Primary beam model.} In order to properly model the instrument response to the sky
we also need an accurate primary beam model. This is especially important for EoR
power spectrum measurements because the (often bright) near-horizon emission resides
at the edge of the foreground wedge 
\citep{Pober:2016, Thyagarajan:2015, Thyagarajan:2015b}, but the low elevation beam
can be difficult to model at the necessary precision \citep{Thyagarajan:2016}. 
\citet{Asad:2016} demonstrated the need for a polarized primary beam model in order
to mitigate the risk of diffuse polarized emission leaking into Stokes I, conflating with a 
potential cosmological measurement. The MWA
primary beam model is continuously improving to incorporate mutual coupling between the 
phased dipoles, and embedded element patterns for each of the 32 dipoles on each antenna 
\citep{Sutinjo:2015}. Other methods are also being pursued to directly measure the full beam 
response in situ using the ORBCOMM satellite constellation \citep{Neben:2015}, and 
drone-flown calibrator sources (\textbf{D.~C.}~Jacobs, et al. in prep). The latter method is similar to
the strategy \citet{Virone:2014} used to verify SKA beam patterns.
